% ============================================================
% VII. DISCUSSION
% ============================================================
\section{Discussion}
\label{sec:discussion}

% ---- A. Practical Trade-offs ----
\subsection{When to Use Each Mode}

The three approximation modes serve different use cases in the robotics pipeline.
\textbf{Convex mode} produces the geometrically tightest approximation and is the best choice when the target simulator or collision checker handles convex meshes natively.
It requires no parametric decomposition and preserves the mesh's convex envelope faithfully.
However, it remains a mesh representation and does not admit the constant-time distance evaluation of analytic primitives.

\textbf{Sphere mode} is well-suited to pipelines that exploit sphere algebra---GPU-based collision pipelines, signed-distance fields via sphere-sums, or planners that can use spherical bounding volume hierarchies.
The medial tree produces multiple small spheres that can nestle into concave regions, often giving tighter overall coverage than capsules on shapes with deep concavities.
The trade-off is higher primitive count: a typical sphere-tree approximation requires many more primitives than the equivalent capsule fit.

\textbf{Capsule mode} is the best choice when smooth, low-count analytic primitives are desired.
Capsules provide constant-time point-to-capsule distance, making them attractive for collision-atlas precomputation, swept-volume computation along trajectories, and analytical signed-distance fields for motion planning.
The principal-axis sectioning assumption means that capsules are most effective on elongated, roughly tubular link geometries---precisely the shape of most robot arm links.

% ---- B. URDF Compatibility ----
\subsection{URDF Compatibility Constraint}

A significant practical constraint is the absence of a native capsule element in the URDF specification.
Our toolchain must decompose each capsule into a cylinder plus two spheres for the collision URDF, which introduces a subtle issue: a simulator loading the URDF reconstructs the collision shape as three separate collision primitives rather than as a single capsule.
This means the simulator does not benefit from the capsule's analytic distance function.
The JSON sidecar addresses this limitation by storing the canonical capsule parameters separately, but bridging the gap between URDF-compatible emission and analytic usability requires downstream support for reading the sidecar.

The broader implication is that the robotics community would benefit from extending URDF (or its successor formats) to include a standard capsule primitive.
Until such an extension is adopted, toolchains like ours must maintain dual outputs.

% ---- C. Coverage vs. Tightness ----
\subsection{Coverage vs. Tightness Trade-off}

The requirement that every mesh vertex be covered by at least one capsule (conservative coverage) creates an inherent tension with tightness.
A single capsule covering an entire link is always guaranteed to exist (the bounding capsule), but it may have significant excess volume.
Splitting the link into multiple capsules reduces the excess volume at the cost of increased primitive count.

Our splitting strategy---guided by the $r/\mathrm{binMed}$ ratio and $\mathrm{capV/aabb}$ ceiling---addresses this trade-off by selectively splitting only those capsules where the radius inflation indicates geometric inefficiency.
This avoids over-splitting in regions where a single capsule is already tight.

% ---- D. PCA Assumption ----
\subsection{Limitation of PCA Sectioning}

The principal-axis sectioning approach makes a strong assumption about link geometry: that each link has a clear longest axis along which capsules should be oriented.
While this holds for standard manipulator links, it may perform poorly on highly branched, flat, or spherical geometry.
For such links, alternative axis selection strategies (e.g., fitting multiple axis orientations or using a volumetric decomposition before capsule fitting) could improve results.
Additionally, the PCA axis captures only the global elongation; locally curved links may benefit from curved capsule chains or articulated capsule hierarchies.

% ---- E. Reproducibility ----
\subsection{Reproducibility and Practical Use}

The toolchain emphasizes reproducible execution through deterministic random seeds, fixed preset definitions, and Docker packaging.
The built-in validation gates allow users to define acceptance criteria for approximation quality (e.g., ``$\mathrm{capV/aabb} < 2.5$ for all links'') and check them programmatically via the \texttt{validate} and \texttt{compare} CLI commands.

For users who need to tune beyond the named presets, the YAML configuration system provides fine-grained control over all fitting parameters.
The recommended workflow is to start with the \texttt{default} preset, inspect the metrics via \texttt{validate}, adjust parameters iteratively, and use \texttt{compare} to verify that the candidate preset improves upon the baseline.
