% ============================================================
% VI. EXPERIMENTS
% ============================================================
\section{Experiments}
\label{sec:experiments}

This section presents an experimental evaluation of the toolchain on the Franka FR3 robot arm.

% ---- A. Research Questions ----
\subsection{Research Questions}

We define five research questions that guide the experimental design:

\begin{itemize}
    \item \textbf{RQ1 (Coverage):} Does the generated geometry preserve coverage of all robot link meshes across all modes and presets?
    \item \textbf{RQ2 (Mode comparison):} How do convex, sphere, and capsule approximation modes trade off primitive count and geometric tightness on a representative robot model?
    \item \textbf{RQ3 (Capsule ablation):} How sensitive is capsule quality to section count, capsule budget, adaptive circle fitting, and splitting thresholds?
    \item \textbf{RQ4 (Runtime):} What computational cost does the full URDF approximation pipeline impose for each mode and preset?
    \item \textbf{RQ5 (Mesh source):} How do visual-mesh and collision-mesh sources affect the fitting quality and metrics?
\end{itemize}

% ---- B. Dataset ----
\subsection{Dataset}

The primary test platform is the Franka FR3 robot arm \cite{franka2020fr3}, consisting of 7 arm links (link0--link7) plus hand and finger links.
The URDF model is bundled in the repository under \texttt{resources/fr3/} and includes both visual meshes (COLLADA \texttt{.dae} files representing the true link geometry) and collision meshes (pre-simplified STL files).
We evaluate on the 7 arm links, which contain the richest geometry variation.
Robots with greater mesh complexity or more diverse link shapes (e.g., quadrupeds, mobile manipulators) may exhibit different trade-offs and are left for future work.

% ---- C. Baselines ----
\subsection{Baselines and Configurations}

For each mode, we evaluate the named presets listed in Table~\ref{tab:presets}.

\begin{table}[t]
\centering
\caption{Modes, presets, and key configuration parameters evaluated in the experiments.}
\label{tab:presets}
\begin{tabular}{@{}llccc@{}}
\toprule
Mode & Preset & Sections & Max Capsules & Adaptive \\
\midrule
Convex & default & -- & -- & -- \\
Sphere & single & -- & 1 per link & -- \\
Sphere & default & -- & -- & -- \\
Capsule & single & 2 & 1 per link & No \\
Capsule & default & 4 & 12 per link & No \\
Capsule & high\_detail & 6 & 16 per link & No \\
\bottomrule
\end{tabular}
\end{table}

For capsule ablations, we additionally evaluate:
\begin{itemize}
    \item Fixed vs.\ adaptive circle count (\texttt{AdaptiveCircleCount} true/false).
    \item \texttt{NSections} values of 2, 4, 6, 8.
    \item \texttt{MaxCapsulesPerLink} values of 1, 4, 8, 12, 16.
    \item Splitting enabled (\texttt{MaxRadiusBinRatio}=1.45) vs.\ disabled (\texttt{MaxRadiusBinRatio}=-1).
    \item Visual mesh vs.\ collision mesh source.
\end{itemize}

% ---- D. Experimental Setup ----
\subsection{Experimental Setup}

All experiments are run on a single machine with an Intel Core i7-13700K CPU, 32~GB RAM, and an NVIDIA RTX 4090 GPU (used only for PyBullet visualization, not for computation).
The pipeline is invoked via Docker (\texttt{urdfapprox:1.5.0}), using the Python API with \texttt{mesh\_source="visual"} unless otherwise noted.
Runtimes are reported as wall-clock time averaged over three runs and include Docker container startup overhead (approximately 1--2~s per invocation).
The capsule union volume is estimated with \texttt{UnionVolumeSamplesPerAxis}=64 during validation, while generation uses the default 32 for speed.

% ---- E. Mode Comparison ----
\subsection{Mode Comparison}

Table~\ref{tab:mode_comparison} presents the aggregate results for all six preset-mode combinations on the FR3 arm.
Per-link data and full experiment logs are available in the artifact directory.

\begin{table}[t]
\centering
\caption{Aggregate mode comparison on the Franka FR3 arm (7 links: link0--link7). Worst-case metrics report the maximum (worst) value across all arm links.}
\label{tab:mode_comparison}
\begin{tabular}{@{}lcrrrr@{}}
\toprule
Mode & Preset & Primitives & Worst capV/aabb & Worst $r$/binMed & Runtime (s) \\
\midrule
Convex & default & 8 & -- & -- & 0.5 \\
Sphere & single & 8 & -- & -- & 10.1 \\
Sphere & default & 63 & -- & -- & 40.7 \\
Capsule & single & 8 & 1.37 & 1.27 & 12.7 \\
Capsule & default & 10 & 1.47 & 1.50 & 33.8 \\
Capsule & high\_detail & 10 & 1.32 & 1.32 & 21.8 \\
\bottomrule
\multicolumn{6}{@{}l}{\small \textit{Note.} Dash (---) indicates metric not applicable (capV/aabb and $r$/binMed are capsule-specific).}\\
\end{tabular}
\end{table}

\textbf{Convex mode} produces one convex hull per link (8 primitives total) and is the fastest configuration at 0.5~s.
Because the convex hull encloses the original mesh vertices by construction, coverage is guaranteed.
However, the output is mesh-based and does not admit the closed-form distance evaluation available for analytic primitives.

\textbf{Sphere-tree modes} trade primitive count against coverage tightness.
The \texttt{single} preset produces one sphere per link (8 total) with near-perfect coverage: only two links exhibit sub-micrometer uncovered vertices ($2\times 10^{-5}$~m worst-case).
The \texttt{default} preset generates 63 spheres across the 7 links but leaves all links with small uncovered regions (worst-case $1.8\times 10^{-2}$~m).
This behavior is by design --- the sphere tree is a medial-axis approximation that places spheres to match the shape's interior, not to enclose all surface vertices.

\textbf{Capsule modes} output 8--10 primitives with tightness quantified by the union-volume ratio \texttt{capV/aabb} (worst-case range 1.32--1.47 across presets) and the radius-inflation metric \texttt{r/binMed} (worst-case range 1.27--1.50).
The \texttt{high\_detail} preset achieves the best worst-case uncovered distance (2.3~mm) and a slightly lower \texttt{capV/aabb} (1.32 vs.\ 1.47 for default), while keeping the total primitive count at 10 with a slightly different per-link distribution.
Runtime for capsule modes ranges from 12.7~s (single) to 33.8~s (default), with the \texttt{high\_detail} preset completing in 21.8~s.

A key finding is that \textbf{no analytic-primitive mode achieves complete (all-vertex) coverage on all arm links}.
Only the convex mesh output guarantees full enclosure.
Analytic primitives inevitably involve trade-offs between tightness and coverage, and the optimal configuration depends on the downstream application's tolerance for small uncovered regions.

% ---- F. Capsule Ablation ----
\subsection{Capsule Ablation}

Table~\ref{tab:capsule_ablation} reports the effect of individual capsule-fitting parameters.
Each row varies one parameter from the default configuration (\texttt{NSections}=4, \texttt{MaxCapsulesPerLink}=12, \texttt{AdaptiveCircleCount}=false, \texttt{MaxRadiusBinRatio}=1.45, \texttt{mesh\_source}=visual) while holding others fixed.

\begin{table}[t]
\centering
\caption{Capsule ablation on the FR3 arm. Each row varies one parameter from the default configuration. Worst-case metrics are across all 7 arm links.}
\label{tab:capsule_ablation}
\begin{tabular}{@{}lccrrrr@{}}
\toprule
Variant & Parameter change & Caps & Worst Dist (mm) & capV/aabb & r/binMed & Time (s)\\
\midrule
nsections\_2 & NSections=2 & 8 & 6.69 & 1.37 & 1.27 & 13.8 \\
nsections\_4 & (default) & 10 & 10.23 & 1.47 & 1.50 & 33.0 \\
nsections\_6 & NSections=6 & 10 & 2.34 & 1.32 & 1.32 & 21.0 \\
nsections\_8 & NSections=8 & 11 & 1.20 & 1.56 & 1.46 & 21.2 \\
\midrule
mcpl\_1 & MaxCapsulesPerLink=1 & 8 & 10.21 & 1.33 & 1.31 & 14.7 \\
mcpl\_4 & MaxCapsulesPerLink=4 & 10 & 10.23 & 1.47 & 1.50 & 32.6 \\
mcpl\_8 & MaxCapsulesPerLink=8 & 10 & 10.23 & 1.47 & 1.50 & 32.4 \\
mcpl\_16 & MaxCapsulesPerLink=16 & 10 & 10.23 & 1.47 & 1.50 & 32.5 \\
\midrule
adaptive\_true & AdaptiveCircleCount=true & 39 & 1.35 & 1.41 & 2.05 & 56.8 \\
\midrule
mrbr\_disabled & MaxRadiusBinRatio=-1 & 10 & 10.23 & 1.47 & 1.50 & 12.6 \\
\midrule
mesh\_collision & mesh source=collision & 10 & 10.23 & 1.47 & 1.50 & 32.5 \\
\bottomrule
\end{tabular}
\end{table}

\textbf{NSections is the dominant control parameter.}
Increasing the section count from 2 to 8 reduces the worst-case uncovered distance from 6.69~mm to 1.20~mm (5.6$\times$ improvement) at the cost of raising the capsule count from 8 to 11 and increasing \texttt{capV/aabb} from 1.37 to 1.56.
However, the relationship is not monotonic: the default configuration (NSections=4) exhibits a \emph{worse} worst-case uncovered distance (10.23~mm) than NSections=2 (6.69~mm), because the 4-section fit splits link0 into two capsules, creating a coverage gap at the split boundary.
At NSections=8, this gap nearly disappears (9~$\mu$m residual on link0).

\textbf{MaxCapsulesPerLink plateaus.}
Values above 4 all produce identical output (10 capsules total, same metrics), indicating that the split logic's improvement threshold (\texttt{MinSplitVolumeImprovement}=0.005) rejects additional capsules as not beneficial regardless of budget.
The budget is therefore not binding at the default value of 12 on this robot.

\textbf{AdaptiveCircleCount} enables per-section multi-circle fitting, which dramatically increases the capsule count from 10 to 39 (3.9$\times$) and runtime from 33~s to 57~s.
The worst-case uncovered distance improves from 10.23~mm to 1.35~mm (7.6$\times$), but the radius-inflation metric \texttt{r/binMed} spikes to 2.05 on link2, indicating significant axial radius variation within individual capsules.

\textbf{MaxRadiusBinRatio} set to --1 (disabling all splitting) produces metrics identical to the default configuration in which the ratio is active at 1.45.
This confirms that the other pipeline mechanisms (COA threshold, \texttt{MinSplitVolumeImprovement}) already prevent excessive radius variation before \texttt{MaxRadiusBinRatio} becomes the active constraint, at least at this threshold value.

\textbf{Mesh source} (visual vs.\ collision) produces identical capsule fits on the FR3, suggesting that the collision meshes (simplified convex envelopes) capture the same geometric structure as the visual meshes for this robot.
This finding may not generalize to robots with greater divergence between visual and collision mesh representations.

\textbf{No configuration achieves full coverage on all arm links.}
Every variant leaves at least one link with a positive worst-case signed distance, confirming that the coverage-tightness trade-off is fundamental to the capsule representation.

% ---- G. Summary ----
\subsection{Summary of Findings}

The experiments yield the following key findings:
\begin{enumerate}
    \item Convex mode is fastest (0.5~s) and guarantees coverage, but produces only mesh-based output.
    \item Sphere-tree modes provide 8--63 primitives; the single-sphere preset offers near-perfect coverage at minimal primitive count but the default tree does not guarantee enclosure.
    \item Capsule modes deliver 8--11 primitives with quantifiable tightness (\texttt{capV/aabb} 1.32--1.56, \texttt{r/binMed} 1.27--1.50).
    \item \texttt{NSections} is the dominant capsule-tuning parameter, improving worst-case coverage from 6.7~mm to 1.2~mm as section count increases from 2 to 8.
    \item \texttt{AdaptiveCircleCount} dramatically increases capsule count (3.9$\times$) but introduces radius inflation.
    \item No analytic-primitive mode achieves full coverage on all links --- this is a fundamental limitation of the representation.
\end{enumerate}

All experimental data, configuration files, and generated figures are stored in the artifact directories corresponding to each experiment request.
