% ============================================================
% I. INTRODUCTION
% ============================================================
\section{Introduction}
\label{sec:intro}

% ---- 1. Problem context ----
Robot simulation, motion planning, and control all depend on efficient collision geometry.
A robot's collision model must be lightweight enough for real-time queries yet faithful enough to avoid missed collisions or excessive false positives.
In the Robot Operating System (ROS) ecosystem, the Unified Robot Description Format (URDF)~\cite{urdf_doc} provides the standard representation, supporting both analytic primitives (spheres, cylinders, boxes) and triangular meshes.
In practice, robot descriptions commonly carry high-detail visual meshes alongside simplified collision meshes or manually authored approximations, leading to representations that are either computationally expensive or inconsistently maintained.

% ---- 2. Existing practice -%
Current practice for collision geometry authoring is largely manual.
A robot designer may hand-place spheres or capsules inside a link mesh, simplify an existing mesh with a 3D modeling tool, or rely on a convex-hull decomposition pipeline.
These approaches are time-consuming, hard to reproduce across different robots and simulators, and often produce collision models at an arbitrary level of detail.
Several simulators and planning frameworks---notably MuJoCo~\cite{todorov2012mujoco}, Drake~\cite{drake}, and PyBullet~\cite{coumans2018pybullet}---support primitive collision geometries that admit closed-form distance and contact computations, but the generation of such primitives from existing URDF descriptions remains a gap.

% ---- 3. Analytic primitives -%
Analytic collision primitives offer significant advantages over mesh-based representations.
Spheres, capsules, and boxes support stable signed-distance evaluation, analytical contact gradients, and efficient broad-phase culling in collision pipelines~\cite{pan2012fcl}.
Capsules (swept spheres) are particularly well suited to robot links because their cylindrical body with hemispherical ends matches the elongated, rounded shape of many manipulator links.
A capsule's distance function---the point-to-segment distance minus the capsule radius---can be evaluated in constant time, making it attractive for real-time collision checking and for computing swept volumes along motion trajectories.
Despite these advantages, no widely adopted tool exists to automatically convert URDF mesh geometry into capsule-based collision primitives with provable coverage guarantees.

% ---- 4. Our approach -%
In this paper we introduce URDFApproxGeom, an integrated open-source toolchain that reads a URDF with mesh collision (or visual) geometry and outputs drop-in replacement URDFs with convex, sphere-tree, or capsule collision approximations.
The toolchain is available as a Python package with a CLI, a public Python API, Docker support, and bundled robot resources.
Our primary methodological contribution is an automatic capsule-fitting pipeline that proceeds in five stages:
(i)~computing the principal axis of each link mesh via PCA;
(ii)~slicing the mesh with planes perpendicular to this axis and extracting 2D cross-section contours;
(iii)~fitting covering circles to each cross section using minimum enclosing circles and, where needed, adaptive Lloyd-style refinement to handle concave or multi-lobed sections;
(iv)~chaining matched circles across adjacent sections into capsules, followed by collinear merging, coverage-preserving radius growth, endpoint tightening, inflation-based splitting, and nested-capsule pruning;
(v)~emitting each capsule as a URDF-compatible cylinder-plus-two-sphere decomposition alongside a JSON sidecar with the canonical analytic parameters.

% ---- 5. Broader system -%
Beyond capsule fitting, the toolchain provides two additional approximation modes for completeness: convex hull generation via CGAL/libigl and sphere-tree approximation via a medial-axis tree algorithm.
All modes emit standard URDF and are accompanied by a CLI validation utility that reports coverage, volume tightness, and radius-inflation metrics.
A built-in comparison workflow generates side-by-side visualizations across all modes and presets, enabling qualitative and quantitative inspection without external tools.

% ---- 6. Contributions -%
In summary, this paper contributes:
\begin{itemize}
    \item \textbf{An integrated URDF-to-primitive approximation pipeline} that preserves visual, inertial, and joint structure while replacing collision geometry with convex, sphere, or capsule primitives.
    \item \textbf{A capsule-fitting method} that produces coverage-preserving, low-count analytic capsule primitives from robot link meshes through PCA sectioning, adaptive circle fitting, and tightness-guided refinement.
    \item \textbf{A reproducible CLI/Python workflow} with named presets, validation metrics, pass/fail gates, and visualization bundles for qualitative comparison.
    \item \textbf{An experimental protocol} and reference results on the Franka FR3 robot (Section~\ref{sec:experiments}) for evaluating coverage, volume tightness, primitive count, runtime, and preset sensitivity, including a capsule ablation study that characterizes the sensitivity of each fitting parameter.
\end{itemize}
